Optimization of respiratory syncytial virus prophylaxis using a generalized additive model of Connecticut hospitalizations

### 2018-01-14 Thesis/Thesis.rnw
More plot refactoring, added bibliography

<<load>>=
read_chunk("../Analysis/RSV.R")
read_chunk("../Thesis/Thesis.R")
@

<<external>>=
@

@@

@@

In children, however, the infection often incites more serious respiratory illness, sometimes leading to hospitalization.

@@

Immunoprophylaxis with palivizumab, a monoclonal antibody medication, has been approved for prevention of RSV infection since 1998. Palivizumab is administered via intramuscular injection, and requires monthly dosing. Its price is \$1500-\$3000 per dose.

Palivizumab has been found to be effective in reducing hospitalizations due to RSV; in double-blinded trials in high-risk infants, it reduced hospitalization rate by 45-55\% compared to placebo.

Owing to the high cost, the American Academy of Pediatrics (AAP) recommends palivizumab immunoprophylaxis only for high-risk infants, and only during RSV season. These guidelines were most recently revised in 2014.

In their guidelines, the AAP acknowledges the potential significance of spatial and temporal variation in RSV incidence, but does not recommend local variance from nation-wide prophylaxis guidelines outside of Florida.

@@

The Department of Public Health of the state of Connecticut, via its Office of Health Care Access (OHCA), maintains a database of all hospitalizations in every non-federal short-stay acute-care general hospital in the state. This database has previously been used to show regional variability in time of peak RSV incidence within Connecticut.

@@

    tikz("figure/seasonByRegion.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(data) +
        theme_light(base_size=10) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=county), width=.1) + 
        geom_point(aes(y=median, x=county)) + 
        labs(aes(x=variable)) +
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        labs(y="Surveillance week") + 
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
        )
    dev.off()
@
\begin{center}
\input{figure/seasonByRegion.tex}
\caption{Onset and offset of RSV season in Connecticut by region}
\label{fig:seasonByRegion}
\end{center}
\end{figure}

@@

\begin{figure}
<<coverageByRegionAAP, results='hide'>>=
    tikz("figure/coverageByRegionAAP.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(unprotectedByCounty %>% filter(strat=="aap") %>% filter(county %in% countiesForPlots)) +
        theme_light(base_size=10) +
        geom_errorbar(aes(ymax=frac.upper, ymin=frac.lower, x=county), width=.1) + 
        geom_point(aes(y=frac.median, x=county)) + 
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        labs(y="Preventable fraction") +
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank()
        )
    dev.off()
@
\input{figure/coverageByRegionAAP.tex}
\caption{Fraction of RSV cases occurring in Connecticut during the protection window of palivizumab prophylaxis administered according to AAP guidelines, by region.}
\label{fig:coverageByRegionAAP}
\end{figure}

@@

\begin{figure}
<<coverageByRegion, results='hide', warning=FALSE>>=
    data = unprotectedByCounty %>% 
        filter(county %in% countiesForPlots)
    tikz("figure/coverageByRegionRegimen.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(data) +
        theme_light(base_size=10) +
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        scale_linetype_manual(values=c(1, 3, 4, 2, 3, 4, 2)) + 
        scale_color_manual(values=c(grey(0), rep(hsv(0, 1, 1), 3), rep(hsv(4/6, 1, 1), 3))) + 
        geom_errorbar(
            aes(ymin=frac.lower, ymax=frac.upper, x=county, linetype=strat, color=strat, group=strat), 
            width=0, 
            position=position_dodge(.5)
        ) + 
        geom_point(aes(y=frac.median, x=county, color=strat), position=position_dodge(.5)) +
        geom_point(aes(y=frac.lower, x=county, color=strat), position=position_dodge(.5), shape=18) +
        geom_point(aes(y=frac.upper, x=county, color=strat), position=position_dodge(.5), shape=18) +
        labs(y="Preventable fraction") +
        guides(color=guide_legend(keywidth=3, keyheight=1), linetype=guide_legend(keywidth=3, keyheight=1)) + 
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            legend.position="bottom"
        )
    dev.off()
@
\input{figure/coverageByRegionRegimen.tex}
\caption{Fraction of RSV cases occurring in Connecticut during the protection window of palivizumab prophylaxis, by region and prophylaxis regimen.}
\label{fig:coverageByRegionRegimen}
\end{figure}

@@

    tikz("figure/onsetOffsetByYear.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(data) +
        theme_light(base_size=10) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=year), width=.1) + 
        geom_point(aes(y=median, x=year)) +
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLebeller) +
        scale_x_continuous(breaks=seq(1990, 2020, by=5)) +
        labs(y="Surveillance week") + 
        theme(
            axis.title.x=element_blank()
        )

    dev.off()
@
\input{figure/onsetOffsetByYear.tex}
\caption{RSV season onset and offset by three-year superseason.}
\label{fig:onsetOffsetByYear}
\end{figure}

@@

>>>>

<<load>>=
library(reshape)
library(dplyr)
library(mgcv)
library(outbreakpredict)
library(zipcode)
library(ggplot2)
library(easyGgplot2)
library(data.table)
library(splitstackshape)

read_chunk("../Analysis/RSV.R")
read_chunk("../Thesis/Thesis.R")
@

<<external, cache=TRUE>>=
@

@@

\tableofcontents
\clearpage

\listoftables
\clearpage

\listoffigures
\clearpage

@@

In children, however, the infection often incites more serious respiratory illness, making RSV a leading cause of hospitalization among infants.\cite{Hall:2009gi,Hall:2013jo,Zhou:2012ho,Boyce:2000gz}

@@

Immunoprophylaxis with palivizumab, a monoclonal antibody medication, has been approved for prevention of RSV infection since 1998. Palivizumab is administered via intramuscular injection, and requires monthly dosing.\cite{Andabaka:2013dr} Its price is \$1500-\$3000 per dose.\cite{Hampp:2011ju}

Palivizumab has been found to be effective in reducing hospitalizations due to RSV; in double-blinded trials in high-risk infants, it reduced hospitalization rate by 45-55\% compared to placebo.\cite{Group:1998ih,Feltes:2003wm}

Owing to the high cost, the American Academy of Pediatrics (AAP) recommends palivizumab immunoprophylaxis only for high-risk infants, and only during RSV season.\cite{Anonymous:1998up,AmericanAcademyofPediatricsCommitteeonInfectiousDiseasesandCommitteeonFetusandNewborn:2003ug,AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj} These guidelines were most recently revised in 2014.

In their guidelines, the AAP acknowledges the potential significance of spatial and temporal variation in RSV incidence, but does not recommend local variance from nation-wide prophylaxis guidelines outside of Florida.\cite{AmericanAcademyofPediatricsCommitteeonInfectiousDiseases:2014bj}

@@

The Department of Public Health of the state of Connecticut, via its Office of Health Care Access (OHCA), maintains a database of all hospitalizations in every non-federal short-stay acute-care general hospital in the state. Previous analysis of this database has shown regional variability in time of peak RSV incidence within Connecticut.\cite{Noveroske:y4fi3188}

@@

    tikz("figure/seasonByRegion.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=county), width=.1) + 
        geom_point(aes(y=median, x=county)) + 
        labs(aes(x=variable)) +
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        labs(y="Surveillance week") + 
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
        )
    dev.off()
@
\begin{center}
\input{figure/seasonByRegion.tex}
\caption{Onset and offset of RSV season in Connecticut by region}
\label{fig:seasonByRegion}
\end{center}
\end{figure}

@@

\begin{figure}
<<coverageByRegionAAP, results='hide'>>=
    tikz("figure/coverageByRegionAAP.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(unprotectedByCounty %>% filter(strat=="aap") %>% filter(county %in% countiesForPlots)) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymax=frac.upper, ymin=frac.lower, x=county), width=.1) + 
        geom_point(aes(y=frac.median, x=county)) + 
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        labs(y="Preventable fraction") +
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank()
        )
    dev.off()
@
\input{figure/coverageByRegionAAP.tex}
\caption{Fraction of RSV cases occurring in Connecticut during the protection window of palivizumab prophylaxis administered according to AAP guidelines, by region.}
\label{fig:coverageByRegionAAP}
\end{figure}

@@

\begin{figure}
<<coverageByRegion, results='hide', warning=FALSE>>=
    data = unprotectedByCounty %>% 
        filter(county %in% countiesForPlots)
    tikz("figure/coverageByRegionRegimen.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        scale_y_continuous(labels=latexPercent, limits=c(0.9, 1), breaks=seq(0, 1, by=.02), expand=c(0.05, 0)) +
        scale_linetype_manual(values=c(1, 3, 4, 2, 3, 4, 2)) + 
        scale_color_manual(values=c(grey(0), rep(hsv(0, 1, 1), 3), rep(hsv(4/6, 1, 1), 3))) + 
        geom_errorbar(
            aes(ymin=frac.lower, ymax=frac.upper, x=county, linetype=strat, color=strat, group=strat), 
            width=0, 
            position=position_dodge(.5)
        ) + 
        geom_point(aes(y=frac.median, x=county, color=strat), position=position_dodge(.5)) +
        geom_point(aes(y=frac.lower, x=county, color=strat), position=position_dodge(.5), shape=18) +
        geom_point(aes(y=frac.upper, x=county, color=strat), position=position_dodge(.5), shape=18) +
        labs(y="Preventable fraction") +
        guides(color=guide_legend(keywidth=3, keyheight=1), linetype=guide_legend(keywidth=3, keyheight=1)) + 
        theme(
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            legend.position="bottom"
        )
    dev.off()
@
\input{figure/coverageByRegionRegimen.tex}
\caption{Fraction of RSV cases occurring in Connecticut during the protection window of palivizumab prophylaxis, by region and prophylaxis regimen.}
\label{fig:coverageByRegionRegimen}
\end{figure}

@@

    tikz("figure/onsetOffsetByYear.tex", width=plotWidth, height=plotHeight, pointsize=10)
    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=year), width=.1) + 
        geom_point(aes(y=median, x=year)) +
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        scale_x_continuous(breaks=seq(1990, 2020, by=5)) +
        labs(y="Surveillance week") + 
        theme(
            axis.title.x=element_blank()
        )

    dev.off()
@
\input{figure/onsetOffsetByYear.tex}
\caption{RSV season onset and offset by three-year surseason.}
\label{fig:onsetOffsetByYear}
\end{figure}

@@

\printbibliography[title={Bibliography}] % Print the bibliography, section title in curly brackets

### 2018-01-14 Thesis/structure.tex
More plot refactoring, added bibliography

\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{example.bib} % The filename of the bibliography

>>>>

\usepackage[backend=bibtex,style=numeric,sorting=none,autocite=superscript]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
%\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{Thesis.bib} % The filename of the bibliography
\renewcommand{\cite}[1]{\autocite{#1}}
